---
title: "Initializing reportifyr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Initializing reportifyr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# reportifyr

```{r setup}
library(reportifyr)
```

`reportifyr` loads with a message specifying some options that are used
to set various values for python setup. You can choose to leave these
options unset and default values will be used instead. Or, if you are
comfortable setting these values, you can alter where the python virtual
environment (.venv) is created and which versions of python,
`python-docx`, and `PyYAML` are used.

## Initializing report project and reportifyr

The first thing to do when using `reportifyr` is to call the
`initialize_report_project` function. This does several things - it
creates all needed directories for saving output figures and tables,
sets up the 'report' directory, and copies a `standard_footnotes.yaml`
file into the 'report' directory that can be used to save footnote
definitions. Additionally, it sets up the needed python dependencies in
a virtual environment so that the python portion of the `reportifyr`
package can properly function.

You'll need to provide a directory (typically the project root
directory) where `reportifyr` will look for a few other directories. If
the directories are not found, the `initialize_report_project` function
will create them. Specifically, the 'report' directory, as well as
several sub directories, will be created if not found. The 'OUTPUTS'
directory, containing 'tables', 'figures', and 'listings' sub
directories, will be created if they do not exist. A '.venv' directory
will be created, if it doesn't exist, and `python-docx` and `PyYAML`
will be saved to this directory.

```{r initialize}
options("venv_dir" = file.path(here::here(), "vignettes"))
initialize_report_project(file.path(here::here(), "vignettes"))
```

When the virtual environment is initially created (i.e., the first time
you call `initialize_report_project`), `reportifyr` will notify you of
which versions python, `python-docx`, and `PyYAML` are being used and
where the virtual environment is created. These versions can be set with
options before running this function and are displayed when `reportifyr`
is loaded.

Here are all the directories and files that exist after running
`initialize_report_project`:

```{r file_outputs, echo = FALSE}
for (dir in list.dirs(file.path(here::here(), "vignettes"), recursive = FALSE)) {
  if (basename(dir) != ".venv") {
    for (subdir in list.files(dir)) {
      cat(paste0(gsub(".*reportifyr/", "reportifyr/", dir), "/", subdir))
      cat("\n")
    }
  } else {
    cat(gsub(".*reportifyr/", "reportifyr/", dir))
    cat("\n")
  }
}
```

```{r remove directories, include = FALSE}
unlink(file.path(here::here(), "vignettes", ".venv"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "OUTPUTS"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "report"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "logfile.log"))
```
