---
title: "Initializing reportifyr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Initializing reportifyr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# reportifyr

```{r setup}
library(reportifyr)
```

On first interaction, `reportifyr` loads with a message specifying options for python setup. These options may be altered to designate where the python virtual
environment (.venv) is created and which versions of python,
`python-docx`, and `PyYAML` are used. Or in most cases, these options may be left unset and default values will be used.

## Initializing report project and reportifyr

The first step in starting up `reportifyrp` is to call the
`initialize_report_project` function. This one function performs several tasks - it creates all needed directories for saving output figures and tables,
sets up the 'report' directory, and copies a `standard_footnotes.yaml`
file into the 'report' directory that can be used to save footnote
definitions. Additionally, it configures the required Python dependencies within a virtual environment to ensure the python components of the `reportifyr` package function correctly.

A directory must be specified in the `initialize_report_project` function (typically the project root directory) where `reportifyr` will search for certain subdirectories. If these subdirectories are not found, the `initialize_report_project` function will create them automacially. The subdirectories that should be found or created are: 

*'report' 
*'OUTPUTS' (containing 'tables', 'figures', and 'listings') 
*'.venv' (containing `python-docx` and `PyYAML`)


```{r initialize}
options("venv_dir" = file.path(here::here(), "vignettes"))
initialize_report_project(file.path(here::here(), "vignettes"))
```

#> When the virtual environment is initially created (i.e., the first time `initialize_report_project` is called), `reportifyr` will notify you of which versions python, `python-docx`, and `PyYAML` are being used and where the virtual environment is created. These versions can be set with options before running this function and are displayed when `reportifyr` is loaded.
#> i feel like this section is redundant and may be confusing if interpreted as a new step

Below is an example of all the directories and files that should exist after running
`initialize_report_project` sucessfully:

```{r file_outputs, echo = FALSE}
for (dir in list.dirs(file.path(here::here(), "vignettes"), recursive = FALSE)) {
  if (basename(dir) != ".venv") {
    for (subdir in list.files(dir)) {
      cat(paste0(gsub(".*reportifyr/", "reportifyr/", dir), "/", subdir))
      cat("\n")
    }
  } else {
    cat(gsub(".*reportifyr/", "reportifyr/", dir))
    cat("\n")
  }
}
```

```{r remove directories, include = FALSE}
unlink(file.path(here::here(), "vignettes", ".venv"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "OUTPUTS"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "report"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "logfile.log"))
```
