---
title: "Integrating reportifyr into analyses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Integrating reportifyr into analyses}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

# Initializing reportifyr

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(reportifyr)

options("venv_dir" = file.path(here::here(), "vignettes"))
initialize_report_project(file.path(here::here(), "vignettes"))
```

If you'd like more detail on the initialization of `reportifyr`, please
see the [vignette on initializing
reportifyr](initializing_reportifyr.html).

# Generating a figure for use with reportifyr

Let's first create a concentration-time plot of the Theophylline
dataset:

```{r theophylline, message=FALSE}
library(ggplot2)
library(dplyr)

data <- Theoph
p <- ggplot(data, aes(x = Time, y = conc, group = Subject)) +
  geom_point() + 
  geom_line() +
  theme_bw() +
  labs(x = "Time (hr)", y = "Theophylline concentration (mg/L)")

p
```

## Saving the figure and creating metadata

If we want to include this plot into the report we'll need to save this
figure to the 'OUTPUTS/figures' directory. There are two separate
processes to accomplish this - the easiest way is to call the wrapper
function `ggsave_with_metadata`:

```{r saving plots with metadata}
figures_path <- file.path(here::here(), "vignettes", "OUTPUTS", "figures")
plot_file_name <- "theoph-pk-plot.png"

ggsave_with_metadata(
  filename = file.path(figures_path, plot_file_name), 
  plot = p,
  width = 6,
  height = 4
)
```

Alternatively you could call `ggplot2::ggsave` and then call
`write_object_metadata`:

```{r saving plots}
ggplot2::ggsave(filename = file.path(figures_path, plot_file_name), 
                plot = p, 
                width = 6,
                heigh = 4)
write_object_metadata(object_file = file.path(figures_path, plot_file_name))
```

Both processes give the same final result, however, severing the tie
between saving an object and writing its metadata may lead to the
these objects becoming out of sync.

This brings us to metadata. `reportifyr` uses metadata to create a
record of the object being inserted into the word document to aid in
reproducibility.

`reportifyr` also uses a meta_type parameter to inject standard
footnotes into the word document (more on that in the [vignette on
integrating reportifyr into report
writing](Integrating_reportifyr_into_report_writing.html)). We can use
the function `get_meta_type` to see what figure and/or table meta types
are currently saved in the `standard_footnotes.yaml` within the 'report'
directory upon initialization. We just need to provide the path to the
`standard_footnotes.yaml` we want to pull meta_types from.

```{r meta types}
meta_types = get_meta_type(file.path(here::here(), "vignettes", "report", "standard_footnotes.yaml"))
names(meta_types)
```

Now, let's re-save the plot object with the `conc-time-trajectories`
meta_type:

```{r meta type update}
figures_path <- file.path(here::here(), "vignettes", "OUTPUTS", "figures")
meta_plot_file_name <- "theoph-pk-plot-mt.png"

ggsave_with_metadata(
  filename = file.path(figures_path, meta_plot_file_name), 
  meta_type = meta_types$`conc-time-trajectories`,
  width = 6,
  height = 4
)
```

Using meta_types also allows for tab-completion to help you get the
meta_type correct! Let's use the `preview_metadata_files` to see the
difference when using a meta type:

```{r}
metadata <- preview_metadata_files(file_dir = figures_path)
knitr::kable(metadata)
```

There is also the `get_meta_abbrevs` function to see standard
abbreviations that can be added:

```{r}
meta_abbrevs <- get_meta_abbrevs(
  file.path(here::here(), "vignettes", "report", "standard_footnotes.yaml")
)
names(meta_abbrevs)
```

Let's perform a simple analysis computing subject level pharmacokinetic
parameters and then study level statistics of the data set to generate
an additional plot. We'll compute maximum drug concentration (Cmax),
time to peak drug concentration (Tmax), and area under the
concentration-time curve (AUC) for the Theophylline data set before
performing a linear regression between weight and AUC.

```{r}
calc_auc_linear_log <- function(time, conc) {
  auc <- 0
  
  cmax_index <- which.max(conc)
  
  for (i in 1:(length(time) - 1)) {
    delta_t <- time[i + 1] - time[i]
    
    if (i < cmax_index) {
      
      auc <- auc + delta_t * (conc[i + 1] + conc[i]) / 2
    } else if (i >= cmax_index && conc[i + 1] > 0 && conc[i] > 0) {
      
      auc <- auc + delta_t * (conc[i] - conc[i + 1]) / log(conc[i] / conc[i + 1])
    } else {
      
      auc <- auc + delta_t * (conc[i + 1] + conc[i]) / 2
    }
  }
  
  return(auc)
}

pk_params <- data %>%
  mutate(Subject = as.numeric(Subject)) %>% 
  group_by(Subject) %>%
  summarise(
    cmax = max(conc, na.rm = TRUE), 
    tmax = Time[which.max(conc)], 
    auc = calc_auc_linear_log(Time, conc),
    wt = Wt %>% unique()
  )

knitr::kable(pk_params)
```

```{r}
lr <- pk_params %>% 
  ggplot(aes(x = wt, y = auc)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "blue") +
  theme_bw() +
  labs(x = "Subject weight (kg)", y = "AUC (hr mg/L)")

lr
```

Let's save this plot with the logistic-regression-plot meta type:

```{r save linear regression}
plot_file_name <- "theoph-pk-exposure.png"
ggsave_with_metadata(filename = file.path(figures_path, plot_file_name), 
                     meta_type = meta_types$`linear-regression-plot`,
                     plot = lr,
                     width = 6,
                     height = 4)
```

And let's view all the metadata files again:

```{r metadata preview}
new_meta <- preview_metadata_files(figures_path)
knitr::kable(new_meta)
```

# Generating a table for use with reportifyr

Now, let's save out the pk_params data frame to .csv so we can include
it in the report. We can pass in any argument that would be used in
`write.csv` as well. Let's use `row.names = FALSE`:

```{r, table saving, message=FALSE}
tables_path <- file.path(here::here(), "vignettes", "OUTPUTS", "tables")
outfile_name <- "theoph-pk-parameters.csv"

write_csv_with_metadata(
  pk_params, 
  file = file.path(tables_path, outfile_name),
  row.names = FALSE
)
```

Let's also save out the Theophylline data set to include in the report.
We can save this as an .RDS file, just to highlight the functionality:

```{r}
data_outfile_name <- "theoph-pk-data.RDS"
save_rds_with_metadata(data, file = file.path(tables_path, data_outfile_name))
```

At the end, we’ve generated two figures and two tables and saved them
alongside their metadata using `reportifyr`, allowing us to easily
incorporate them into our upcoming report writing!

```{r remove directories, include = FALSE}
unlink(file.path(here::here(), "vignettes", ".venv"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "OUTPUTS"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "report"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "logfile.log"))
```
