---
title: "Integrating reportifyr into report writing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Integrating reportifyr into report writing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(reportifyr)

options("venv_dir" = file.path(here::here(), "vignettes"))
initialize_report_project(file.path(here::here(), "vignettes"))
```

```{r, include = FALSE, echo=FALSE, message=FALSE, warning=FALSE, output=FALSE}
library(ggplot2)
library(dplyr)

result <- file.copy(from = system.file("extdata/template.docx", package = "reportifyr"),
                    to = file.path(here::here(), "vignettes", "report", "shell", "template.docx"))

data <- Theoph
p <- ggplot(data, aes(x = Time, y = conc, group = Subject)) +
  geom_point() + 
  geom_line() +
  theme_bw() +
  labs(x = "Time (hr)", y = "Theophylline concentration (mg/L)")

meta_types = get_meta_type(file.path(here::here(), "vignettes", "report", "standard_footnotes.yaml"))

figures_path <- file.path(here::here(), "vignettes", "OUTPUTS", "figures")
meta_plot_file_name <- "theoph-pk-plot.png"

ggsave_with_metadata(
  filename = file.path(figures_path, meta_plot_file_name), 
  meta_type = meta_types$`conc-time-trajectories`,
  width = 6,
  height = 3
)

calc_auc_linear_log <- function(time, conc) {
  auc <- 0
  
  cmax_index <- which.max(conc)
  
  for (i in 1:(length(time) - 1)) {
    delta_t <- time[i + 1] - time[i]
    
    if (i < cmax_index) {
      
      auc <- auc + delta_t * (conc[i + 1] + conc[i]) / 2
    } else if (i >= cmax_index && conc[i + 1] > 0 && conc[i] > 0) {
      
      auc <- auc + delta_t * (conc[i] - conc[i + 1]) / log(conc[i] / conc[i + 1])
    } else {
      
      auc <- auc + delta_t * (conc[i + 1] + conc[i]) / 2
    }
  }
  
  return(auc)
}

pk_params <- data %>%
  mutate(Subject = as.numeric(Subject)) %>% 
  group_by(Subject) %>%
  summarise(
    cmax = max(conc, na.rm = TRUE), 
    tmax = Time[which.max(conc)], 
    auc = calc_auc_linear_log(Time, conc),
    wt = Wt %>% unique()
  )

lr <- pk_params %>% 
  ggplot(aes(x = wt, y = auc)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "blue") +
  theme_bw() +
  labs(x = "Subject weight (kg)", y = "AUC (hr mg/L)")

plot_file_name <- "theoph-pk-exposure.png"
ggsave_with_metadata(
  filename = file.path(figures_path, plot_file_name),
  meta_type = meta_types$`linear-regression-plot`,
  plot = lr,
  width = 6,
  height = 3
)

tables_path <- file.path(here::here(), "vignettes", "OUTPUTS", "tables")
outfile_name <- "theoph-pk-parameters.csv"

write_csv_with_metadata(
  pk_params, 
  file = file.path(tables_path, outfile_name),
  row.names = FALSE
)

data_outfile_name <- "theoph-pk-data.RDS"
save_rds_with_metadata(data, file = file.path(tables_path, data_outfile_name))
```

If you'd like more detail on the initialization of `reportifyr`, please see the [vignette on initializing reportifyr](initializing_reportifyr.html).

To build reports with `reportifyr` you need to insert a figure or table caption using the 'Insert Caption' button under the 'References' ribbon.

![](data/images/insert_caption.avif){width=100%}

This will insert a caption (either figure or table) into the Microsoft Word document. Underneath that caption, you need to add:

{rpfy}:file_name_of_figure/table

This string triggers the magic behind `reportifyr`, and inserts the figure or table correctly into the document. The `file_name_of_figure/table` will need to be in the appropriate 'OUTPUTS/' directory with the associated metadata file.

In the [vignette on integrating reportifyr into analyses](Integrating_reportifyr_into_analyses.html), we saved two figures and two tables. In the template Microsoft Word document packaged with `reportifyr`, figure and table captions, and their associated magic strings to allow insertion of figures/tables and footnotes, have already been inserted. Below is an image of the caption and magic string for the concentration time profile figure we generated:

![](data/images/figure-2_caption_and_magic_string.avif){width=100%}

We can now perform some setup and start filling in the template! The first thing we'll do is assign the input and output document names. We will be populating the `template.docx` which contains no figures or tables inserted by `reportifyr`. If the input document does contain `reportifyr` inserted objects, you'll need to remove them with `remove_tables_figures_footnotes` first.

```{r, table list}
docx_shell <- file.path(here::here(), "vignettes", "report", "shell", "template.docx")
doc_dirs <- make_doc_dirs(docx_in = docx_shell)
```

The `make_doc_dirs` helper function creates several file paths that are useful for this process:

```{r}
doc_dirs
```

If your shell contains `reportifyr` figures and tables, we can clean those by running the following function:

```{r, remove reportifyr content}
remove_tables_figures_footnotes(docx_in = docx_shell,
                                docx_out = doc_dirs$doc_clean)
```

This will save a clean draft in the 'report/draft' directory ready for the next processing steps.

We start by inserting the tables

```{r, tables}
tables_path <- file.path(here::here(), "vignettes", "OUTPUTS", "tables")

add_tables(docx_in = doc_dirs$doc_clean,
           docx_out = doc_dirs$doc_tables,
           tables_path = tables_path)
```

We can see the tables were successfully inserted:

![](data/images/table_2_in_doc.avif){width=100%}

Now that the tables are inserted, we can insert the figures:

```{r, figures}
figures_path <- file.path(here::here(), "vignettes", "OUTPUTS", "figures")

add_plots(docx_in = doc_dirs$doc_tables,
          docx_out = doc_dirs$doc_tabs_figs,
          figures_path = figures_path)
```

And now we can see our figures in the document as well:

![](data/images/figure_2_in_doc.avif){width=100%}

Now we can add the footnotes:

```{r, footnotes}
footnotes <- file.path(here::here(), "vignettes", "report", "standard_footnotes.yaml")

add_footnotes(docx_in = doc_dirs$doc_tabs_figs,
              docx_out = doc_dirs$doc_draft,
              figures_path = figures_path,
              tables_path = tables_path,
              footnotes = footnotes)
```

We now can see our figures and tables have footnotes inserted:

![](data/images/fig_2_with_fn.avif){width=100%}
![](data/images/table_2_with_fn.avif){width=100%}

Now that all the elements are inserted, we can generate a clean final report. This severs the link between `reportifyr` and the document, so please be careful using this function:

```{r}
final_report <- file.path(here::here(), "vignettes", "report", "final", "report.docx")

finalize_document(docx_in = doc_dirs$doc_draft,
                     docx_out = final_report)
```

This creates a finalized document, free of magic strings, while also capturing the metadata for this final document. The metadata includes a file hash so you can see if this document has been updated since its creation.

![](data/images/fig_2_final.avif){width=100%}
![](data/images/table_2_final.avif){width=100%}

```{r remove directories, include = FALSE}
unlink(file.path(here::here(), "vignettes", ".venv"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "OUTPUTS"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "report"), recursive = TRUE)
unlink(file.path(here::here(), "vignettes", "logfile.log"))
```
